Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-07-29 review-responses.tex

  %%%%%%%%%%%%%%
  \reviewersection{2}
  \begin{quote}
    \color{gray}
    In this manuscript, Gower et al. extend the functionality of stdpopsim to include selection. 
    They test the performance of a few different methods that infer historical population sizes, 
    the distribution of fitness effects of new deleterious mutations, and those that detect selective sweeps using the simulations in SLiM. 
    The added functionality will be useful to the users of stdpopsim.
\end{quote}

  Thanks for your constructive comments on our paper.

  \begin{point}{\revref} % \&75
    How do you deal with overlapping annotations in a gff?
  \end{point}

  \reply{
    We have added a note to the implementation section about how we handle overlapping annotations in a GFF file to \revref.
    In short, annotations can't overlap (we remove overlapping bits);
    see the documentation at \url{https://popsim-consortium.github.io/stdpopsim-docs/stable/api.html#stdpopsim.Contig.add_dfe}.
  }

### 2025-07-29 analysis2_manuscript.tex
Apply suggestions from code review
Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    %IG: I moved this mention of h-s relationship up to here with a forward ref to newly added methods section
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep[][see \textbf{Methods}]{huber2018gene}.
    %by defining ranges of selection coefficients and associating each range with a different dominance coefficient. 
    %IG: I moved this label to the newly added methods section --> \revpoint{1}{5}
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. 
    For the purposes of \stdpopsim, we do not allow for overlapping annotations -- 
    genomic elements of the given annotation being modeled (e.g., CDS) are prioritized over overlapping annotations of other types (e.g., exons)
    and we do not currently support multiple overlapping annotations of different types. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).

>>>>

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    %IG: I moved this mention of h-s relationship up to here with a forward ref to newly added methods section
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep[][see \textbf{Methods}]{huber2018gene}.
    %by defining ranges of selection coefficients and associating each range with a different dominance coefficient. 
    %IG: I moved this label to the newly added methods section --> \revpoint{1}{5}
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. 
    For the purposes of \stdpopsim, we do not allow for overlapping annotations -- 
    any interval covered by an annotation that is added to a given contig
    is removed from any previously-added annotations. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
